---
title: "Statistical Inference - Part1"
author: "Shimon Yannay"
date: "3/12/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Statistical Inference Course Project
## Overview
The project consists of two parts:

A simulation exercise.
Basic inferential data analysis.

## Part 1: Simulation Exercise Instructions
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

### Show the sample mean and compare it to the theoretical mean of the distribution
```{r 1.1, echo=TRUE, warning=FALSE}
lambda <- 0.2
n <- 40
num_sim <- 1000
# Generate 40 samples X 1000 times
set.seed(9876)
exp <- matrix(rexp(n * num_sim, rate = lambda), num_sim)

# Calculate the average of each row
expMeans <- rowMeans(exp)

# Compare the means of the simulation and the theory
simMean <- mean(expMeans)
simMean

# Calculate the Mean by theory
theoryMean <- 1/lambda
theoryMean

```
### Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution
```{r 1.2, echo=TRUE, warning=FALSE}
# Variance of simulation
expVar <- var(expMeans)
expVar

# Variance according to theory
theorySD <- 1/(lambda * sqrt(n))
theoryVar <- (theorySD^2)/n
theoryVar

# Simulation SD
simSD <- sd(expMeans)
simSD
```
### Show that the distribution is approximately normal
```{r 1.3, echo=TRUE, warning=FALSE}
plotdata <- data.frame(expMeans)
g <- ggplot(plotdata, aes(x =expMeans))
g <- g + geom_histogram(aes(y=..density..), color="black", fill = "lightgray", binwidth = 0.25)
g <- g + labs(title = "Distribution of averages of 40 Samples", x = "Mean", y = "Density")
g <- g + geom_vline(aes(xintercept = simMean, color = "red"))
g <- g + geom_vline(aes(xintercept = theoryMean, color = "green"))
g <- g + stat_function(fun = dnorm, args = list(mean = simMean, sd = simSD), color = "blue", size = 1.0)
g <- g + stat_function(fun = dnorm, args = list(mean = theoryMean, sd = theorySD), color = "purple", size = 1.0)
g
```
### Conclusions
The density of the actual data is shown by the light gray bars. The theoretical mean and the sample mean are very close.The blue line shows the normal curve formed by the the theoretical mean and standard deviation. The purple line shows the curve formed by the sample mean and sample standard deviation.

Thus, the distribution of averages of the 40 exponential distributions is very close to the normal distribution