---
title: "Statistical Inference - Part2"
author: "Shimon Yannay"
date: "3/12/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Statistical Inference Course Project
## Overview
The project consists of two parts:

A simulation exercise.
Basic inferential data analysis.


## Part 2: Basic Inferential Data Analysis Instructionsless 
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

### Load the ToothGrowth data and perform some basic exploratory data analyses

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Usage
ToothGrowth
Format
A data frame with 60 observations on 3 variables.

[,1]	len	numeric	Tooth length
[,2]	supp	factor	Supplement type (VC or OJ).
[,3]	dose	numeric	Dose in milligrams/day

```{r 2.1, echo=TRUE}
data("ToothGrowth")
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
ggplot(aes(x=dose, y = len), data = ToothGrowth) + 
     labs(x = "Length", y = "Frequency", title = "Tooth Length per Dose and Supplement") +
    geom_point(aes(color = supp))

ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity") +
    facet_grid(. ~ supp) +
    xlab("Dose") +
    ylab("Length") +
    labs(title = "Tooth Length per Dose") +
    guides(fill=guide_legend(title="Supplement"))

ggplot(data=ToothGrowth, aes(x=as.factor(supp), y=len, fill=dose)) +
    geom_bar(stat="identity") +
    facet_grid(. ~ dose) +
    xlab("Supplement") +
    ylab("Length") +
    labs(title = "Tooth Length per Supplement") +
    guides(fill=guide_legend(title="Dose"))
```

### Provide a basic summary of the data

We can see that there is an impact on tooth growth with regard to two parameters: 1. By increasing the dose value 2. By supplementing the vitamins with OJ rather than with VC

```{r 2.2, echo=TRUE}

```

### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
```{r 2.3, echo=TRUE}
t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"], ToothGrowth$len[ToothGrowth$supp=="VC"], paired = FALSE, var.equal = FALSE)
```
The p-value is 0.06063 and the confidence interval is [-0.1710156 7.5710156] (containing the 0)

Since p-value is 0.06063 (bigger than 0.05), it is not satisfactory to reject the null hypothesis.
We cannot assume the supplement type has a significant effect on tooth growth

```{r}
t.test(ToothGrowth$len[ToothGrowth$dose==2], ToothGrowth$len[ToothGrowth$dose==1], paired = FALSE, var.equal = TRUE)
```
We see the p-value is very small, and is significant. Therefore, we can reject the null hypothesis and assume the dose increase from 1mg to 2mg has a positive effect on tooth growth
```{r}
t.test(ToothGrowth$len[ToothGrowth$dose==1], ToothGrowth$len[ToothGrowth$dose==0.5], paired = FALSE, var.equal = TRUE)
```
We see that the p-value is small, meaning it is significant.

This concludes tha we can reject the null hypothesis and assume the dose increase from 0.5mg to 1mg has a positive effect on tooth growth


### State your conclusions and the assumptions needed for your conclusions
#### Assumptions 
- Guinea pigs have been sampled randomly.
- Environmental influence of both groups are the same

#### Conclusions
- The supplement type does not show a significant increase in tooth growth even though it does have a confidence level that crosses 0 at the 95% confidence.

- It appears that there an increase in tooth growth when the dose is increased
